Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters

### 2010-08-24 doc/paper/main.tex
simulation results about done, need to change the activation graph.

The timing of the single flashes were:
1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350,
61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063,
90.169, 96.994, 110.644, 117.469, 120.881, 130.988, 132.694, 154.875,
158.288, 161.700, 165.113, 168.525, 176.925, 178.631, 183.750, 190.575,
204.225, 205.931, 211.050, 216.038, 222.863, 226.275, 236.513, 248.456,
255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394,
299.381

Timing of the double flashes were:
0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756,
91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000,
134.400, 136.106, 141.225, 144.638, 156.581, 159.994, 166.819, 171.806,
175.219, 185.456, 188.869, 202.519, 212.756, 217.744, 221.156, 227.981,
229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506,
272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500

@@

%\begin{table}[t]
%\centering
%\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
%\hline
%$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & Residual & Error\\
%\hline
%\rowcolor[gray]{.8}
%1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
%\hline
%\hline
%1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
%1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
%1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
%1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
%1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
%1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
%1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
%1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
%1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
%1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
%1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
%\hline
%1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
%\hline
%\end{tabular}
%\caption{Estimated Parameters on 11 separate 310.8 second runs. 
%    True value in first row,
%    and average in last. Residual measured as $\sqrt{MSE}$ with noisy signal,
%    Error measured as $\sqrt{MSE}$ with true (noiseless) signal.}
%\label{tab:LowNoiseResults}
%\end{table}

The minimum and maximum error particles had MSE of $ $ and $ $,
respectively, while the mean of the particles gave MSE of $ $. 
%TODO fill in errors
The correlation matrix for the final parameter estimates is shown in 
\autoref{tab:long_corr}.
The significant correlation of parameters indicate that no one estimate of 
particles dominate the solution. Additionally, given the very low error in 
the BOLD,
the energy graph is full of global minima.

The final histogram for a single active region, 
with a single set of true parameters is shown in \autoref{fig:slicesim_hist2}.
This also indicates that ill-defined nature of the parameters.

%parameters:
%&1.151&  0.353& 0.380& 0.026& 1.98&  2.333& 1.645 \\
\begin{figure*}[!t] %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
\caption{Histogram of estimated parameters. Average SNR in the region was 
$0.97$ and Mutual Information between simulation and BOLD estimate was greater 
than $0.15$. TODO: add actual error.}
\label{fig:slicesim_hist2}
\end{figure*}

\begin{figure}[!t]
\subfloat[Region labels for simulated slice.]
{\includegraphics[width=2.5in]{images/simregions.png}\label{fig:simslice_mask}}
\hfil
\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
{\label{fig:sim_hm_mi} \includegraphics[width=2.5in]{images/sim_hm_mi}}
\caption{Comparison of activation with greymatter, parameter regions.}
\label{fig:simslice_hm}
\end{figure}

\begin{itemize}
\item Long Single Result: Correlation Matrix, error in estimate
\item Possum Results: Histogram of parameters for a region
\item FMRI Results: Activation maps comparing M.I. with SPM
\end{itemize}

>>>>

The timing of the single flashes were:
\small{1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350, 
61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063, 
90.169, 96.994, 110.644, 117.469, 120.881, 130.988, 132.694, 154.875, 
158.288, 161.700, 165.113, 168.525, 176.925, 178.631, 183.750, 190.575, 
204.225, 205.931, 211.050, 216.038, 222.863, 226.275, 236.513, 248.456, 
255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394, 
299.381}

Timing of the double flashes were:
\small{ 0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756, 
91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000, 
134.400, 136.106, 141.225, 144.638, 156.581, 159.994, 166.819, 171.806, 
175.219, 185.456, 188.869, 202.519, 212.756, 217.744, 221.156, 227.981, 
229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506, 
272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500 }

@@

%TODO double check the error measurement, add min/max/mean errors for the runs
\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\centering
\begin{tabular}{|c | c | c | c | c | c | c | }
\hline
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
\hline
\rowcolor[gray]{.8}
1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  \\
\hline
\hline
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 \\
\hline
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 \\
\hline
\end{tabular}
\caption{Estimated Parameters on 11 separate 310.8 second runs. 
    True value in first row,
    and average in last.}
\label{tab:LowNoiseResults}
\end{table}

%TODO fill in errors
The final square root Mean-Squared-Error (between the estimated and 
noiseless BOLD signal) for the 2400s run was %TODO
and the final correlation matrix is shown in \autoref{tab:long_corr}.

The final parameter estimates of the 310.8s runs are shown in
\autoref{tab:LowNoiseResults}; the square-root MSE for these
parameters ranged from $0.002833$ to $0.003345$, with a mean of 
$0.003066$. Peak signals reached levels of $0.025$.

\begin{figure}[!t]
\subfloat[Region labels for simulated slice.]
{\includegraphics[width=3cm]{images/simregions.png}\label{fig:simslice_mask}}
\hfil
\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
{\label{fig:sim_hm_mi} \includegraphics[width=4cm]{images/sim_hm_15_6.png}}
\subfigure{\includegraphics[width=.6cm]{images/scale6.png}}
\caption{Comparison of activation with greymatter, parameter regions.}
\label{fig:simslice_hm}
\end{figure}

%todo fix these histograms with marker for true value and drop A1,A2
\begin{figure*}[!t] %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
\caption{Histogram of estimated parameters. Average SNR in the region was 
$0.97$ and Mutual Information between simulation and BOLD estimate was greater 
than $0.15$.}
\label{fig:slicesim_hist2}
\end{figure*}

\begin{figure*}[!t] %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
\caption{Histogram of estimated parameters for region 3. Average SNR in the 
region was 
$0.39$ and Mutual Information between simulation and BOLD estimate was greater 
than $0.15$.}
\label{fig:slicesim_hist3}
\end{figure*}

In the POSSUM simulation, areas passing a threshold of
mutual information of $0.15$  are shown in 
\autoref{fig:sim_hm_mi}
whereas the active region numbers are shown in 
\autoref{fig:simslice_mask}.
Note that region 4 had a very small $\epsilon$ and so
had an extremely low $SNR$ ($<.1$). The histograms of parameter
estimates for sections
2 and 3 are shown in \autoref{fig:slicesim_hist2} and 
\autoref{fig:slicesim_hist3}, respectively. Region 1 is
left out because of the low number of points and because
its SNR was between region 2 and region 3 ($0.8$), so it
was a less interesting result.

\subsection{Real Data}
\begin{figure*}
\centering
\subfigure[]{\label{fig:hm_spm} \includegraphics[scale=.85]{images/spm_hm}}
\subfigure[]{\label{fig:hm_canon_spm_x} \includegraphics[scale=3]{images/spm_hm_x}}
\subfigure[]{\label{fig:hm_canon_spm_y} \includegraphics[scale=3]{images/spm_hm_y}}
\subfigure[]{\label{fig:hm_canon_spm_z} \includegraphics[scale=3]{images/spm_hm_z}}
\subfigure{\label{fig:scale_spm} \includegraphics[scale=.5]{images/scale1}}
\caption{SPM results. Units of activation are in Student's $t$-scores; higher indicates higher
        assurance that the signal cannot have occurred through noise alone.
        Sagittal, coronal and axial slices  are
        \autoref{fig:hm_canon_spm_x}, \autoref{fig:hm_canon_spm_y}, and
         \autoref{fig:hm_canon_spm_z}, respectively. A series of axial slices are
         shown in \autoref{fig:hm_spm}. }
\label{fig:hm_canon_spm}
\end{figure*}
